HUME: Human UCCA-Based Evaluation of Machine Translation

### 2016-03-16 paper/main.tex
ondrej's comments and fixes

Other approaches have been proposed for semantic representation, which are grounded in the words
and phrases of the text, and have generally represented semantic structures as bi-lexical dependencies.
Examples include the tectogrammatical layer \parcite{sgallhp:1986} used in the Prague dependency
treebanks, e.g. the Czech-English one \parcite{hajic2012announcing}, as well
as dependencies derived from Minimal Recursion Semantics representations \parcite{oepen2006discriminant}.
Nevertheless, despite their appeal as deep and intuitive means for semantic representation,
these approaches require linguistic expertise for their annotatation, due to the intricate and
fine-grained distinctions they provide, and are thus in our view less suitable for basing
an interpretable MT evaluation on. 
%\newcite{Basile:12} attempted to make some aspects of the semantic annotation more accessible,
%but to our knowledge no results of this brand have been reported.

@@

Human annotation of the semantic units requires first deciding whether
a unit is an {\it atomic units}, i.e., that it has no meaning-bearing
sub-units, or a {\it structural unit}. In most cases, atomic units would
correspond to individual words, but they may also correspond to opaque
multi-word expressions, i.e., expressions that do not have meaningful sub-units.
When a multi-word unit is labeled as atomic, all its sub-units are discarded
from the evaluation.

@@

Structural units have sub-units (children in the UCCA graph), which may either be
atomic or structural, and represent the relation between the sub-units.
For instance, a Scene unit 
represents the relation between the main relation, participants and secondary
relations in it. Structural units may be labeld as ``Adequate'' or ``Bad'', meaning
that the relation between the sub-units went wrong.

@@

\begin{compactitem}
\item The components of the translation are ordered differently from the source,
  such that the relation between them changes. For instance, where
  the sentence ``man bites dog'' is translated into a sentence meaning ``dog bites man'',
  the semantic unit corresponding to the Scene will be marked as Bad, while
  the inidividual atomic units ``dog'', ``bites'' and ``man'' will be marked as Green.
\item A superfluous word or phrase has been inserted into
  the translation. For instance, if ``dog bites man'' is translated into ``dog visciously bites man''.
\item One of the sub-units is missing from the translation. For instance, if ``dog bites man'' is translated into ``dog bites''.
\end{compactitem}

\XXX{Could someone review this for Ondrej:}
Note that for non-configuration languages, the relationship between the
governing and the dependent units is often expressed using morphological
properties of the dependents instead of their ordering (this holds especially for the verb and its
modifiers, less so for components of noun phrases). Since errors in morphology
should be expressed using the traffic lights on the dependent nodes, we would expect
a higher proportion of ``Adequate'' structural units for Czech and Polish.

>>>>

Other approaches have been proposed for semantic representation, which are grounded in the words
and phrases of the text, and have generally represented semantic structures as bi-lexical dependencies.
Examples include the tectogrammatical layer \parcite{sgallhp:1986} used in the Prague dependency
treebanks, e.g. the Czech-English one \parcite{hajic2012announcing}, as well
as dependencies derived from Minimal Recursion Semantics representations \parcite{oepen2006discriminant}.
Nevertheless, despite their appeal as deep and intuitive means for semantic representation,
these approaches require linguistic expertise for their annotation, due to the intricate and
fine-grained distinctions they provide, and are thus in our view less suitable for basing
an interpretable MT evaluation on. 
%\newcite{Basile:12} attempted to make some aspects of the semantic annotation more accessible,
%but to our knowledge no results of this brand have been reported.

@@

Human annotation of the semantic units requires first deciding whether
a unit is an {\it atomic unit}, i.e., that it has no meaning-bearing
sub-units, or a {\it structural unit}. In most cases, atomic units would
correspond to individual words, but they may also correspond to opaque
multi-word expressions, i.e., expressions that do not have meaningful sub-units.
When a multi-word unit is labeled as atomic, all its sub-units' annotations are discarded
from the evaluation.

@@

Structural units have sub-units (children in the UCCA graph), which may either be
atomic or structural, and represent the relation between the sub-units.
For instance, a Scene unit 
represents the relation between the main relation, participants and secondary
relations in it\XXX{again, could we avoid the word ``relation'' here? Would
``event'' be an acceptable synonym?}. Structural units are labeled as ``Adequate'' or ``Bad'', meaning
that the relation between the sub-units went wrong.

@@

\begin{compactitem}
\item The components of the translation are ordered differently from the source,
  such that the relation between them changes. For instance, where
  the sentence ``man bites dog'' is translated into a sentence meaning ``dog bites man'',
  the semantic unit corresponding to the Scene will be marked as Bad, while
  the individual atomic units ``dog'', ``bites'' and ``man'' will be marked as Green.
\item A superfluous word or phrase has been inserted into
  the translation. For instance, if ``dog bites man'' is translated into ``dog visciously bites man''.
\item One of the sub-units is missing from the translation. For instance, if ``dog bites man'' is translated into ``dog bites''.
\end{compactitem}

\XXX{Could someone review this for Ondrej:}
Note that for non-configurational languages, the relationship between the
governing and the dependent units is often expressed using morphological
properties of the dependents instead of their ordering (this holds especially for the verb and its
modifiers, less so for components of noun phrases). Since errors in morphology
should be expressed using the traffic lights on the dependent nodes, we would expect
a higher proportion of ``Adequate'' structural units for Czech and Polish.
